Speeding up eQTL scans in the BXD population using GPU

### 2019-07-30 paper.tex
Rob edited abstract of paper.tex online with Bitbucket

\affil[$\ast$]{Department of Preventive Medicine, University of Tennessee Health Science Center, Memphis, TN} 
\affil[$\dagger$]{Department of Genetics, Genomics, and Informatics, University of Tennessee Health Science Center, Memphis, TN} 
\affil[$\S$]{Department of Biostatistics, University of Wisconsin-Madison, Madison, WI}

@@

\begin{abstract}
The BXD strains are an important reference population of
recombinant inbred lines which have been phenotyped extensively.
To facilitate interactive use of genotype-phenotype relationships
in this population, we sought to speed up eQTL scans where we
perform a univariate genome scan for every trait in a collection
of omic traits.  By using easily parallelizable operations such as
matrix multiplication, vectorized operations, and elementwise
operations, we are able to decrease runtimes approaching real-time
computation.  We used parallelization using different CPU threads
as well as GPUs.  We found that the speed advantage of GPUs is
dependent on problem size and shape.  These results indicate a pathway for
speeding up eQTL scans using LMMs.  Our implementation is in the
Julia programming language.
\end{abstract}

>>>>

\affil[$\ast$]{Department of Preventive Medicine, University of Tennessee Health Science Center, Memphis, TN} 
\affil[$\dagger$]{Department of Genetics, Genomics and Informatics, University of Tennessee Health Science Center, Memphis, TN} 
\affil[$\S$]{Department of Biostatistics, University of Wisconsin-Madison, Madison, WI}

@@

\begin{abstract}
The BXD strains of mice are an important reference population for
systems biology and genetics that have been full sequenced and deeply phenotyped.
To facilitate interactive use of genotype-phenotype relationships
in this family, we sought to develop new algorithms and code to enable
high performance whole genome QTL scan for massive omics data sets of up to 1 million 
traits. By using easily parallelizable operations including
matrix multiplication, vectorized operations, and elementwise
operations, we have been able to decrease runtimes approaching real-time
computation, making this code ideal for interactive web services, such as
GeneNetwork.org. We used parallelization of different CPU threads
as well as GPUs. We found that the speed advantage of GPUs is
dependent on problem size and shape (n of cases, n of genotypes, n of traits).  
Our results provide a clear path for speeding up eQTL scans using LMMs. 
Our implementation is in the Julia programming language.
\end{abstract}
